---
title: "Tensorflow Contributors"
author: "Augustina Ragwitz"
date: "July 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(readr)
library(reshape2)
library(zoo)
```

# Overview

The goal of this analysis is to discover what Github events can tell us about a project's overall activity level, trends, and growth potential. 

This report serves two purposes. The original inspiration came from a request to better understand how Github was determining the contributor metrics it was reporting on it's website for the Tensorflow project. The second purpose is to test and refine participation metrics from other ongoing research efforts that could identify a high activity, growing project.

This notebook lives in Github and is the second in a series: https://github.com/missaugustina/tensorflow_reports

# Github Events

While commits can provide some insight into a project's overall growth and activity levels, not all contributions come in the form of commits. Github is an excellent resource for tracking this activity and they provide an archive of all public events going back to 2010. This section explores how the Github events activity for the tensorflow repository indicates a growing project.

The main page of the repo has a Contributors number but it's not clear how that is calculated. Their 2016 State of the Octoverse report (https://octoverse.github.com/) provided a defintion for the numbers in that report as: "unique contributors (users who pushed code, opened or commented on an issue or PR), unique code reviewers (users who commented on the changed files), and most forks".

Contributor Activity Described by Octoverse:

 * PushEvent
 * PullRequestReviewCommentEvent
 * PullRequestEvent
 * IssuesEvent
 * IssueCommentEvent
 * ForkEvent
 * CommitCommentEvent
 
 This section will explore these specific Github event types.

## Setting up the Github Events

This section outlines how the datasets were created.

1. Select events from the Github Archive going back to 2015:
https://bigquery.cloud.google.com/savedquery/306220071795:5c2975a4930245e184d1b87c40575821

2. Per Octoverse definition above, select the following event types:
https://bigquery.cloud.google.com/savedquery/306220071795:3c3f978b47324b44a954f026a79556d7

3. For purposes of this study, we are only interested in the 'tensorflow/tensorflow' repository.
https://bigquery.cloud.google.com/savedquery/306220071795:5d53bc77299c4a1195df258a04caa646

4. Export the resulting table to Cloud Storage (GBQ API will only return a limited dataset). Make public and get public link.

5. Download, extract, and read into R! (may need to add the ".csv" suffix)
https://storage.googleapis.com/open_source_community_metrics_exports/githubarchive_contributor_events_tf_repo_only.gz

```{r include=FALSE}
tf_gh_events <- read_csv('downloads/githubarchive_contributor_events_tf_repo_only.csv')

tf_gh_events <- tf_gh_events %>% 
  mutate(event_date = as.Date(created_at), event_month = as.yearmon(event_date),
         type=ifelse(type=="PullRequestReviewCommentEvent", "PRRCommentEvent", type))

tf_gh_events_types <- tf_gh_events %>%
  group_by(event_month, type, actor_login) %>% 
  summarise(num_events=n()) %>% 
  mutate(num_events_log=round(log(num_events + 1))) %>%
  group_by(event_month, type) %>%
  mutate(num_actors = n(), 
         num_actors_log=round(log(num_actors + 1)), 
         num_events_type = sum(num_events)) %>% 
  group_by(event_month) %>%
  mutate(total_events=sum(num_events))
```

## Octoverse Contributor Metrics

### Unique Contributors (users who pushed code, opened or commented on an issue or PR)
```{r}
tf_gh_contributors <- tf_gh_events_types %>% 
  filter(
    type == "PushEvent" | # Pushed Code
    type == "IssuesEvent" | type == "IssueCommentEvent" | # Opened or commented on Issue or PR
    type == "PullRequestEvent") %>% # Opened or commented on PR
  group_by(event_month, actor_login) %>%
  summarise(num_events=sum(num_events)) %>%
  group_by(event_month) %>%
  summarise(num_actors=n(), total_events=sum(num_events)) %>%
  mutate(metric="Contributors")

```

### Number of Forks
```{r}
tf_gh_forkers <- tf_gh_events_types %>% 
  filter(type == "ForkEvent") %>%
  group_by(event_month, actor_login) %>%
  summarise(num_events=sum(num_events)) %>%
  group_by(event_month) %>%
  summarise(num_actors=n(), total_events=sum(num_events)) %>%
  mutate(metric="Forkers")
```

### Unique Code Reviewers (users who commented on the changed files)
```{r}
tf_gh_reviewers <- tf_gh_events_types %>% 
  filter(type == "CommitCommentEvent" | type == "PRRCommentEvent") %>%
  group_by(event_month, actor_login) %>%
  summarise(num_events=sum(num_events)) %>%
  group_by(event_month) %>%
  summarise(num_actors=n(), total_events=sum(num_events)) %>%
  mutate(metric="Reviewers")
```

### Summary

Overall there is a trend towards growth in the 3 Actor categories proposed by the Octoverse report. Reviewers show the most growth. Forkers and Contributors appear to follow a very similar pattern. It would be interesting to see if this is indicative of a correlation.

```{r}
tf_gh_octoverse <- rbind(tf_gh_contributors, tf_gh_forkers, tf_gh_reviewers)
rm(tf_gh_forkers, tf_gh_contributors, tf_gh_reviewers)
```

```{r}
ggplot(data = tf_gh_octoverse, 
       aes(x = factor(event_month))) +
  geom_line(aes(colour=metric, y=num_actors, group=metric)) +
  geom_point(aes(colour=metric, y=num_actors, group=metric)) +
  ylab("Actors") +
  xlab("Month") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(data = tf_gh_octoverse, 
       aes(x = factor(event_month))) +
  geom_line(aes(colour=metric, y=num_actors, group=metric)) +
  geom_point(aes(colour=metric, y=num_actors, group=metric)) +
  ylab("Actors (log)") +
  xlab("Month") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_log10()

```


## Looking for Patterns

```{r}
tf_gh_events_type_per_actor <- tf_gh_events %>% 
  group_by(event_month, actor_login, type) %>% 
  summarise(num_actor_events=n()) %>%
  group_by(event_month, actor_login) %>%
  mutate(cnt_types_p_actor=n()) %>%
  group_by(event_month, type, cnt_types_p_actor) %>%
  mutate(num_actors_w_type_cnt=n())
```

```{r}
tf_gh_event_types_events_log <- tf_gh_events_types %>% 
  group_by(event_month, type, num_events_log) %>%
  summarise(actors_cnt=n(), events_min=min(num_events), events_max=max(num_events)) %>%
  group_by(type, num_events_log) %>% 
  mutate(num_events_log_freq = n()) %>%
  group_by(type) %>%
  mutate(max_num_events_log_freq = max(num_events_log_freq)) %>%
  filter(max_num_events_log_freq == num_events_log_freq) %>%
  group_by(type, num_events_log) %>%
  summarise(events_min=min(events_min), 
  events_max=max(events_max), 
  num_events_log_freq=first(num_events_log_freq))

tf_gh_event_types_actors_log <- tf_gh_events_types %>% 
  group_by(event_month, type, num_actors_log) %>%
  summarise(actors_min=min(num_actors), actors_max=max(num_actors)) %>%
  group_by(type, num_actors_log) %>% 
  mutate(num_actors_log_freq = n()) %>%
  group_by(type) %>%
  mutate(max_num_actors_log_freq = max(num_actors_log_freq)) %>%
  filter(max_num_actors_log_freq == num_actors_log_freq) %>%
  group_by(type, num_actors_log) %>%
  summarise(actors_min=min(actors_min), 
  actors_max=max(actors_max), 
  num_actors_log_freq=first(num_actors_log_freq))
  
```

### Event Type Contribution Frequency

What event types have the most unique actors?
What are the most frequent event types (per actor)?

```{r}
ggplot(data = tf_gh_event_types_actors_log, 
       aes(x = factor(num_actors_log), y = num_actors_log_freq, fill=type)) +
  geom_bar(stat="identity", position="dodge") +
  xlab("Actors (log)") +
  ylab("Freq as Max p/ Month") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(data = tf_gh_event_types_events_log, 
       aes(x = factor(num_events_log), y = num_events_log_freq, fill=type)) +
  geom_bar(stat="identity", position="dodge") +
  xlab("Events per Actor (log)") +
  ylab("Freq as Max p/ Month") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


### Event Type Diversity per Actor

How many actors contributed how many different event types for each event type? In other words are some event types more strongly associated with actors who contributed a wider range of event types?

For each month, what event type frequency had the most actors for each type?

```{r}
tf_gh_event_types_summary <- tf_gh_events_type_per_actor %>% 
  group_by(event_month, type) %>% 
  mutate(max_actors = max(num_actors_w_type_cnt)) %>%
  filter(num_actors_w_type_cnt == max_actors) %>%
  summarise(cnt_types_p_actor=first(cnt_types_p_actor),
            max_num_actors_w_type_cnt=first(num_actors_w_type_cnt))

tf_gh_event_types_summary_total <- tf_gh_event_types_summary %>%
  group_by(type, cnt_types_p_actor) %>%
  summarise(cnt_types_p_actor_freq=n()) %>%
  group_by(type) %>% mutate(max_type_freq=max(cnt_types_p_actor_freq))

ggplot(data = tf_gh_event_types_summary_total, 
       aes(x=factor(cnt_types_p_actor), y=cnt_types_p_actor_freq, fill=type)) +
  geom_bar(stat="identity", position="dodge") +
  xlab("Event Types p/ Actor") +
  ylab("Freq of Max p/ Month") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(fill=guide_legend(title="Event Types"))

```

### Events Contributed by Diverse Type Actors

How many events do actors who contribute in the top event type frequencies typically have per type? What's "normal" for active contributors?


# Events vs Commits

## Top Committers Event Types

What event types are strongly correlated with commit proportions per the analysis in this report?
(this will require matching using the Github API)

How about using Github's top committers list?

## Event Type Actors vs Top Committers

How many top actors are missed by only looking at top committers? Are top committers missing from any of the top event types?

